---
title: "AshvilleMaps"
author: "Rash"
date: "19 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
install.packages(ggmap)

library(ggplot2)
library(ggmap)
```


```{r}
ash_data <- read.csv(file.choose(),header = T, sep = ",")
```


#
get_map(location = c(lon = -95.3632715, lat = 29.7632836), zoom = "auto", scale = "auto", maptype = c("terrain", "terrain-background", "satellite", "roadmap", "hybrid", "toner", "watercolor", "terrain-labels", "terrain-lines", "toner-2010", "toner-2011", "toner-background", "toner-hybrid", "toner-labels", "toner-lines", "toner-lite"), source = c("google", "osm", "stamen", "cloudmade"), force = ifelse(source == "google", TRUE, TRUE), messaging = FALSE, urlonly = FALSE, filename = "ggmapTemp", crop = TRUE, color = c("color", "bw"), language = "en-EN", api_key)



```{r}
mapgilbert <- get_map(location = c(lon = mean(ash_data$longitude), lat = mean(ash_data$latitude)), zoom = "auto" ,maptype = "roadmap", source="google",scale = "auto")

ggmap(mapgilbert) +
+     geom_point(data = df, aes(x = ash_data$longitude, y = ash_data$latitude, fill = "red", alpha = 0.8), size = 5, shape = 21) +
+     guides(fill=FALSE, alpha=FALSE, size=FALSE)

```

```{r}

review <- read.csv(file.choose(),header = T,sep = ",")
library(plyr)

merged_reviews <- ddply(review,.(listing_id),
      summarise,
      reviews = paste(unique(review$comments),collapse = ' '))


```

```{r}

dictionary <- read.csv(file.choose(),header = F,sep = "\t")
keys <- dictionary$V1
values <- dictionary$V2
dictionary <- setNames(as.list(values),keys)



```

```{r}
review_comments <- read.csv(file.choose(),header = T,sep = ",")

getScore <- function(text)
{
 words <- strsplit(text," ");
 
 if(length(words)>0)
 {
 words <- unique(words)
 
 score <- 0
 
  for(i in 1:length(words))
  {
    for(j in 1:length(words[[i]])) 
    {
      word <- words[[i]][j]
      
      word <- tolower(word)
    
      if (dictionary[word]=="NULL" || length(word)<1)
      { 
        score = score + 0
      
      } else 
      {
       score <- score + as.double(dictionary[word])   
      }
    }
  
  }
 return(score)
 }else
 {
   return(0)
 }
 
}
 
 
```

```{r}

sentiments <- list()

for (i in 1:nrow(review_comments))
{
  data <- toString(review_comments[i,2])
  
  sentiment <- lapply(data,getScore)
  
  sentiments <- rbind(sentiments,sentiment)
  
  
}





```

```{r}

review_comments <- cbind(review_comments,sentiments)

names(review_comments) <- c("listing_id","Comments","Sentiment_Score")

review_comments$Sentiments <- as.integer(review_comments$Sentiments)

mean(review_comments$Sentiments)

group_mean <- aggregate(review_comments$Sentiment_Score,list(review_comments$listing_id),mean)

write.csv(group_mean,"review_grouped_sentiments.csv")
 

```

```{r}

hist(group$x,col="green",main="Review Sentiment Score",xlab = "Sentiment Score")
abline(v=median(group$x),col="magenta",lwd=4)
abline(v=8,lwd=3)

abline(v=median(group$x),col="magenta",lwd=4)
legend("topright", inset=.05,
       c("Mean","CutOff"), fill=c("magenta","black"), horiz=TRUE)

```


```{r}

 hosts <- read.csv(file.choose(),header = T,sep = ",",stringsAsFactors = FALSE)

hosts$host_since_frm <- as.Date(hosts$host_since,format = "%d-%m-%Y")

hosts$days_since <- (Sys.Date()-hosts$host_since_frm)

hosts$email <- 1
hosts$social <- 1
hosts$phone <- 1
hosts$reviews <- 1

if(length(as.double(grep('facebook',hosts$host_verifications[1])))>0){print(1)} else{print(0)}


checkVerify <- function(column)
{
  for (i in 1:nrow(hosts))
  { flag <- 0
          
    if(length(as.double(grep(column,hosts$host_verifications[i])))>0)
      {flag <- 1} 
    
    hosts[column][i,] <- flag
    
  }
}


```


```{r}

for (i in 1:nrow(hosts))
  { flag <- 0
          
    if(length(as.double(grep("facebook",hosts$host_verifications[i])))>0 ||length(as.double(grep("google",hosts$host_verifications[i])))>0 ||length(as.double(grep("jumio",hosts$host_verifications[i])))>0 ||length(as.double(grep("kba",hosts$host_verifications[i])))>0)
      {flag <- 1} 
    
    hosts[column][i,] <- flag
    
  }


hosts$verification_score <- (hosts$email + hosts$social +hosts$phone +hosts$reviews)


```


```{r}

for (i in 1:nrow(hosts))
  { flag <- 0
   
  if(toString(hosts$host_response_time[i])=="within a few hours")
  {
    flag <- 2
  }else if (toString(hosts$host_response_time[i])=="within an hour")
  {
      flag <- 1
  }else if (toString(hosts$host_response_time[i])=="within a day")
  {
    flag <- 3
  }else if (toString(hosts$host_response_time[i])=="a few days or more")
  {
    flag <- 4
  }
    
    hosts["response_rate"][i,] <- flag
    
  }

```


```{r}
hosts$price <- as.numeric(hosts$price)
model_1 <- lm(hosts$price~1)


scaled_v_score <- scale(hosts$verification_score)
scaled_res_rate <- scale(hosts$response_rate)
scaled_days <- scale(hosts$days_since)
price <- hosts$price


 

```

```{r}

scaled_v_score <- scale(hosts$verification_score)
scaled_res_rate <- scale(hosts$response_rate)
scaled_days <- scale(hosts$days_since)
price <- price[price<50]



```

