---
title: "Employment -Model Selection"
output: pdf_document
date: "21 November 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Employee Data

longley <- read.table(file.choose(),header = TRUE)

```{r longley}
longley <- read.table(file.choose(),header = TRUE);
summary(longley)
```

# Build full model with all the variables and the initial model with no variable

```{r}
model_0 <- lm(y~1,data = longley);
model_full <-lm(y~x1+x2+x3+x4+x5+x6,data = longley);
```

# Run the stepwise forward selection process to identify the best model
```{r}
step(model_0,scope = list(lower=model_0,upper=model_full),direction = "forward");
```

This Analysis shows that with Forward selection process. Model with X2, X3, X4, X6 is the best model with minimum AIC value as shown above. This is the same as we see in the table provided in the book where this model has min RSS of 841173.0036. Hence, the best model would be with constant plus X2, X3, X4 and X6.

# Remove X5 from data set and analyse


```{r}
longley <-longley[,-5]
```

# Regressing X1+X2+X3+X4+X6 on y

```{r}
model <- lm(y~x1+x2+x3+x4+x6, data = longley);
summary(model)
```

# Checking Correlation between variables
```{r}
cor(longley)
```

According to this data we can say that variable X1 is not significant. Rest of the variables look significant here. However, when the correlation is calculated between the variables, it seems that x1 is highly correlated to x2 and x6. This shows that there could be high collinearity in the data.

# Checking VIF for multicollinearity.

```{r}
library("car");
vif(model);
```

This is confirmed by the high VIF values of X1, X2 and X6 as above.

## Including Plots

```{r pressure, echo=TRUE}
par(mfrow=c(1,2))
plot(longley$x6,model$residuals,main="Residual vs Year")
plot(model$fitted.values,model$residuals,main="Residual vs Fitted Value")
 
```

Both the plots look similar and shows non-linear trend in X6(Year) and possible multi-collinearity.

# Possible solution to this could be transforming X6 and apply ridge regression to enforce penalty on variables.

```{r}
library("MASS");
model_ridge<- lm.ridge(y~x1+x2+x3+x4+x6,data= longley, lambda = seq(0,0.1,0.001));
plot(model_ridge);
```

```{r}
select(model_ridge)
```

# We can check the fit of model by using qqplots stud resid vs quantiles
```{r}
par(mfrow=c(1,1))
qqPlot(model)

```

This shows that there are outliers, going out of confidence interval and plot does not show exact linear behaviour. Further model diagnostics can be applied to check outliers in data set, removing influential obs and add/remove variables to get better fit. Analysis can be performed by checking the multicollinearity, regressing with all possible variables and understand further improvements and may be perform various model diagnostics such as overfitting, leverage and cook’s D etc. evaluation along with simulation to get more accurate results.
